---
title: "Natural Disaster Sentiment Analysis"
author: "Scott Burstein"
date: "`r format(Sys.time(), '%d %B, %Y')`"
###"Data Science and Society (Sociology 367) - Final Project"
output: html_document
---

# Setup

## Configure RStudio Rmd File Output Format

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the Libraries and Datasets

### Google Trends Package

[gtrendsR CRAN Package](https://cran.r-project.org/web/packages/gtrendsR/gtrendsR.pdf) 
used to perform and display Google Trends Queries.

Google Trends data is accessible in R through the gTrendsR package, which was 
created by [Philippe Masicotte](https://github.com/PMassicotte) and made 
available through [CRAN (The Comprehensive R Archive Network)](https://cran.r-project.org/).

Google trends data for specified search queries is represented by a `hits` 
variable, which expresses the relative volume of Google searches for specific 
terms over geographical and time parameters.

`hits` - a numeric integer between 0 and 100 representing Google's weekly search
volume as a proportion of the maximum search volume for the specified keyword 
within the given time and location bounds. 
It is worth noting that every Google trends query will *always* have at least 
one interest_over_time `hits` value of 100 for a given query. 

```{r load_packages, message = FALSE}
#---PACKAGES------  > remove "#" if package not previously installed
#install.packages("tidyverse")
library(tidyverse)
#install.packages("gtrendsR")
library(gtrendsR)
#install.packages("ggthemes")
library(ggthemes)
#install.packages("maps")
library(maps)
#install.packages("lubridate")
library(lubridate)
```

### README

See README for more thorough description of the `FEMA_Declarations` data set, 
with content summary, column variable descriptions, acknowledgments, source 
description and licensing information.

### Primary Data Source

[Kaggle Natural Disaster Data set](https://www.kaggle.com/headsortails/us-natural-disaster-declarations). The 
`FEMA_Declarations` data set is used for the bulk of this analysis and is also 
integrated with Google Trends data.

```{r load_data, message = FALSE}
#---DATA--------- > remove "#" to view raw data
FEMA_Declarations <- read_csv("FEMA_data/us_disaster_declarations.csv")
#view(FEMA_Declarations)
```
<br/><br/>

For later analysis, the `small_FEMA` data frame can be used to parse FEMA 
declarations that have occurred over the last 5 years in the states that have, 
in those 5 years, recorded the most FEMA declarations. It is created after 
`recent_FEMA_byState`.

## Filter for Event Types

### Generalizable Parameters for Event Filter

In order to produce research that is both reproducible and timely, certain 
parameters can be initialized once, and then referenced multiple times later. 
Since gTrendsR can easily return Google Trends information from the past five 
years, it is intuitive to create a `date_5y` variable that can be used to filter 
FEMA declarations by date, thus eliminating all those that occurred previously. 
The data set at large contains FEMA declarations from 1953 through the present.
Throughout this research there will also be many variables that are reused. 
In portions of the analysis that copy methods used elsewhere, the assignment 
operator is used to allow for minimal rewriting in the event that certain 
functions are purposed for further analysis in the future.

```{r initialized_parameters, echo=FALSE}
# Start year to filter events (only show FEMA decs. since specified year)
date_5y <- Sys.Date() - 365.25*5

# Number of days after a natural disaster to query "topic" interest for in state
num_days <- 14
```

```{r clean_date, echo=FALSE}
FEMA_Declarations <- FEMA_Declarations %>% 
  # clean date variable
  mutate(declaration_day = gsub( " .*$", "", declaration_date ))
```

---

# Research Questions Preview

**(1)** Do Google search patterns correlate with real-world events?

**(2)** Does interest in climate change increase regionally after the occurrence
of a natural disaster?

**(3)** Do people modify their Google search behavior and become more 
environmentally conscious after a natural disaster?*

---

# Initial Analysis of FEMA Declarations

## All FEMA Declarations by State

```{r FEMA_Decs_by_State, echo=FALSE}
all_FEMA_decs <- FEMA_Declarations %>% 
  # group declarations by state and date
  group_by(state, declaration_day) %>% 
  count() %>% 
  ungroup()
```

`alltime_FEMA_byState` contains the number of FEMA declarations for every US 
state and territory since 1953.

```{r all_FEMA_Declarations_by_State, echo=FALSE}
alltime_FEMA_byState <- all_FEMA_decs %>%
  group_by(state) %>% 
  count() %>% 
  arrange(desc(n))
# data frame w/ 59 states/territories, n count(FEMA declarations)
alltime_FEMA_byState
```

Since the database was created in 1953, the following 10 states have had the 
most FEMA natural disaster emergency declarations called (as of `r Sys.Date()`):

CA, TX, OK, WA, FL, OR, NY, AZ, LA, NM

## Recent FEMA Declarations by State

The date this file was last compiled: 
Current date (`Sys.Date`) is  `r Sys.Date()`.

The date 5 years before, which is the starting time bound for subsequent Google 
Trends queries shown below: 
Variable `date_5y` is `r date_5y`.

`recent_FEMA_byState` contains the number of FEMA declarations for every US 
state and territory from the last 5 years.

```{r recent_FEMA_Decs_by_State, echo=FALSE}
recent_FEMA_byState <- all_FEMA_decs %>%
  filter(declaration_day >= date_5y) %>% 
  group_by(state) %>% 
  count() %>% 
  arrange(desc(n))

recent_FEMA_byState
```

From `r date_5y` to `r Sys.Date()`, the following 10 states have had the 
most FEMA natural disaster emergency declarations called:

CA, WA, OR, OK, FL, TX, NV, AZ, LA, MS

In the past 5 years, California has recorded 75 FEMA declarations, outpacing all 
other states. Much of these can be attributed to the recent spell of wildfires 
which cause massive amounts of damage to property and endanger millions of 
Americans each year.

Florida has issued 23 FEMA declarations in the last 5 years, many of which are 
the result of tropical storms from the Atlantic that grow into hurricanes. 
These also cause billions of dollars in damage, widespread power outages, 
endanger the lives of local residents and also threaten to submerge the already 
eroding eastern coastline of Florida.

```{r small_FEMA_df, echo=FALSE}

n_states <- 10
top_states <- recent_FEMA_byState[["state"]][1:n_states]

date_5y <- Sys.Date() - 365.25*5

small_FEMA <- FEMA_Declarations %>% 
  subset(state %in% top_states) %>%
  mutate(incident_begin_day = gsub( " .*$", "", incident_begin_date ),
            incident_end_day = gsub( " .*$", "", incident_end_date ) ) %>% 
  filter(declaration_day >= date_5y) %>% 
  select(declaration_day, 
         state, 
         designated_area,
         declaration_type,
         incident_type, 
         declaration_title, 
         incident_begin_day, 
         incident_end_day
         )
# add filter() below to specify only statewide FEMA declarations
#%>% filter(designated_area == "Statewide")
#uncomment to view small_FEMA table
#view(small_FEMA)
```

---

# Key Google Trends Summary Insights:

## Climate Change Google Searches by State:

A data frame arranged in descending order of all 50 states and Washington D.C.,
showing the relative hits count for the search term "climate change" over the 
last 5 years.

```{r climate_change_interest_by_state, echo=FALSE}
search_terms <- c(
"climate change"
)

climate_trends <- gtrends(keyword = search_terms,
        geo = "US") # default time is past 5 years

location_df <- data.frame(climate_trends[["interest_by_region"]][["location"]])
hits_df <- data.frame(climate_trends[["interest_by_region"]][["hits"]])

names(location_df) <- "location"
names(hits_df) <- "hits"

climate_location_hits_table <- cbind(location_df$location, hits_df$hits)

climate_location_hits_table <- data.frame(climate_location_hits_table)

names(climate_location_hits_table)[1] <- "location"
names(climate_location_hits_table)[2] <- "hits"

climate_location_hits_table
```

Vermont set the mark for most Google search hits - relative search volume for 
"climate change" over the past five years. Washington D.C. had 0.85 as much 
search volume for the same term. The top ten regions were:

VT, DC, ME, AK, OR, MA, NH, RI, HI, CO

## Climate Change Keywords

Visualizations, such as the faceted line plots below help illustrate important 
trends in sentiment towards climate change and associated topics. For example, 
interest in climate change, global warming, and the green new deal topics all 
spiked during the 2020 election season, as climate change became an important 
topic in the Presidential race.

```{r US_climate_change-top4-query, echo=FALSE, results="Hide"}
#remove results="hide" to view glimpse() of interest_over_time

climate_search_terms <- c(
"climate change",		
"global warming",		
"fossil fuel",		
"green new deal"
)

climate_change_gtrends <- gtrends(keyword = climate_search_terms,
        geo = "US",
        time = "today 12-m") 

climate_change_gtrends[1]$interest_over_time

US_climate_interest <- climate_change_gtrends %>% 
          .$interest_over_time %>%
          glimpse()
```

```{r US_climate_change-viz, echo=FALSE}

 climate_change_facet_plot <- US_climate_interest %>% 
  mutate_at("hits", ~ifelse(. == "<1", 0.5, .)) %>% # replace with 0.5
  mutate_at("hits", ~as.numeric(.)) %>% # convert to numeric
  # Begin ggplot
  ggplot(aes(x = date, y = hits)) +
  geom_line(colour = "darkblue", size = 1.5) +
  facet_wrap(~keyword) +
  ggthemes::theme_economist()

climate_change_facet_plot
```
[Technical Resource Cited](https://martinctc.github.io/blog/vignette-google-trends-with-gtrendsr/)

Sentiment towards climate change can be difficult to track as individuals' 
search queries may contain different keywords. The faceted plot above also shows
that volume for "climate change" searches greatly exceeds the "hits" value of 
synonymous terms, like "global warming" and also tangentially related topics, 
like "fossil fuel" and "green new deal". It is evident that out of the four 
terms plotted above, "climate change" is the most viable as an indicator of the 
general public's sentiment towards climate change. This is likely a result of it 
being the most used term in news outlets and research when referring to global 
changes in weather and the environment.

An interesting property of Google Trends data, which will be proven, is that in
addition to tracking human sentiment towards longstanding topics, like climate 
change, it can also be used to detect one-time events

---

# Research Question 1

**Do Google search patterns correlate with real-world events?**

Causality between what happens in the real-world and Google search trends is a 
prerequisite for subsequent analysis. While it may seem intuitive, it is worth 
explaining the mechanisms by which society and events express themselves in the 
information age. A new adaptation of the proverbial question "does a tree really 
fall in the woods if no one is around to hear it?" may be "do current events 
really occur without being searched for on the internet?". According to existing 
social science research - the short answer is no.

The use of Google's search engine is so ingrained in the United States and 
global cultures, that it is the internet which aggregates and disseminates 
the vast majority of information humans use to interpret the world around them.

In order to illustrate the relationship between Google search patterns and 
the occurrence of natural disasters, it is necessary to view differences in 
interest over time for search queries pertaining to time-specific natural 
disasters. 

Each fall, from August through October, two of the most devastating weather 
trends play out. Both of these phenomena lead to hundreds of deaths and billions
of dollars in damages each year. A suitable method to track social interest in 
these disasters (as they happen) is to plot interest in the disaster name as a 
function of time. If there were to be a relationship, a clear spike in interest 
for the keyword would appear at the time of the event.

## Google Search Correlation with Florida Hurricane Events

The first of these two weather trends is tropical storm formation in the 
Atlantic, which ultimately leads to hurricanes in the southeastern United 
States. As global temperatures rise, the prevalence and intensity of tropical 
storm systems increase as well. Florida is subject to the most intense and 
frequent of these hurricanes.  

```{r FL_hurricane-top4-query, echo=FALSE, results="Hide"}

search_terms <- c(
"hurricane michael",		
"hurricane dorian",		
"hurricane isaias",		
"hurricane sally"
)

FLHurr_gtrends <- gtrends(keyword = search_terms,
        geo = "US-FL",
        time = "today+5-y")  #default "today+5-y" otherwise "today 12-m"

#interest_over_time output can be seen here:
#FLHurr_gtrends[1]$interest_over_time

FL_hurr_interest <- FLHurr_gtrends %>% 
          .$interest_over_time %>% # working .$ by list subset
          glimpse()
```

```{r FL_hurricane-top4-viz, echo=FALSE}
 FL_hurr_plot <- FL_hurr_interest %>% 
  mutate_at("hits", ~ifelse(. == "<1", 0.5, .)) %>% # replace with 0.5
  mutate_at("hits", ~as.numeric(.)) %>% # convert to numeric
  # Begin ggplot
  ggplot(aes(x = date, y = hits)) +
  geom_line(colour = "darkblue", size = 1.5) +
  facet_wrap(~keyword) +
  ggthemes::theme_economist()
FL_hurr_plot
```

It is thus verifiable that all of these events garnered peak interest on Google 
at the precise date that they occurred. The spike in Google query hits is so 
sharp, it seems as though human interest in these events is short-lived and 
mostly isolated at the exact moment of their occurrence. This directly confirms 
the hypothesis of research question 1: Google search volume for a natural 
disaster peaks precisely at the event date.

Hurricane Isaias and Hurricane Sally were two of the most devastating storms to
hit the Gulf Coast and Southeastern United States in 2020. They were classified 
as Category 1 and Category 2 storms respectively, leaving more than $10.0 
billion USD of damage in their wakes.

Conversely, Hurricane Dorian from 2019 and Hurricane Michael from 2018 were both 
Category 5 storms, significantly more powerful and inflicting even more damage. 
Hurricane Dorian is attributed to more than 100 deaths and $5.1 billion USD in 
damage alone. Hurricane Michael caused 74 deaths and led to $25.5 billion USD in
damage. 

Google search interest for these category 5 hurricanes showed no more than 
ripples (hits ≤ 5) for the entirety of this past year. This would be unlikely if
it were the case that event interest on Google was not time dependent; 
Hurricanes Dorian and Michael were significantly more consequential than 
Hurricanes Isaias and Sally.

When the data is analyzed carefully however, it is evident that even search hits
for Hurricanes Dorian and Michael also increased fractionally during the 2020 
hurricane season - around the time Google searches for active hurricanes spiked.
It is perhaps the case that people do become more interested in past natural 
disasters in order to contextualize the events that they are currently 
experiencing.

## Google Search Correlation with California Wildfire Events

The second climate trend is increased temperatures in the western U.S., which 
lead to brush and forest fires in California. In 2018 alone, more than 1,670,000 
acres of land burned, California's most destructive wildfire season to date.

```{r CA_fire-top4-query, echo=FALSE, results="Hide"}
CA_top4_fires <- c(
"bobcat fire",		
"august fire",		
"creek fire",		
"bear fire"
)

CAFire_gtrends <- gtrends(keyword = CA_top4_fires,
        geo = "US",
        time = "today+5-y") #%>% summary()

#interest_over_time output can be seen here:
#CAFire_gtrends[1]$interest_over_time

CA_fire_interest <- CAFire_gtrends %>% 
          .$interest_over_time %>% 
          glimpse()
```

```{r CA_fire-top4-viz, echo=FALSE}
 CA_fire_facet_plot <- CA_fire_interest %>% 
  mutate_at("hits", ~ifelse(. == "<1", 0.5, .)) %>% # replace with 0.5
  mutate_at("hits", ~as.numeric(.)) %>% # convert to numeric
  # Begin ggplot
  ggplot(aes(x = date, y = hits)) +
  geom_line(colour = "dark red", size = 1.5) +
  facet_wrap(~keyword) +
  ggthemes::theme_economist()
CA_fire_facet_plot
```

Like the Google trends output for Hurricanes in Florida, wildfires in California 
show the same property of having peak interest on Google around the time of the 
event itself.

A key difference between wildfires in California and hurricanes in Florida is 
that wildfires in California tend to have a much longer duration, often lasting 
weeks. This is evident by the increased time periods of certain spikes in 
interest, which is even more evident when the `gtrends()` query is performed on 
a shorter duration (such as the past 1 year instead of 5 years). 

Another notable feature of the line plots is that there appears to be multiple 
time periods that return high hits values for a given query. For example, "creek
fire" most notably refers to the largest wildfire in recent California history 
that started on September 4th, 2020 and is still active at the date of 
submission for this paper as of November 22nd, 2020. However, due to the naming 
convention of California wildfires, it also returned Google "hits" values for 
the Creek Fire of greater Los Angeles in 2017 and smaller, similarly named fires
in 2018. This makes it harder to infer when the exact peak in interest for an 
event occurred, as many of the keywords refer to multiple wildfires in recent 
California history. 

An additional example of noise in this search query can be seen in the 
"august fire" graph, where, in addition to a peak at the start of the 2020 
August Complex fire (August 16th, 2020), there are regular jumps visible each 
year in the month of August. This is surely not a coincidence, but rather of 
conflated search queries.

The faceted line plot for interest over time for California wildfires still 
confirms the hypothesis that Google search queries peak at the time of an event, 
as this was still an observable pattern in the data. All four example plots 
returned maximum hit values at days within the duration of each wildfire. 
However, it does prove the need to tread carefully. Depending on the nature of 
weather phenomena, or a weather phenomenon's naming convention, Google search 
queries for associated terms may appear noisy or misleading. Additionally, it 
becomes increasingly challenging to make claims like "local populations tend to/
to not search for past natural disaster events in order to contextualize 
impending events" as the data is more ambiguous.

```{r Generalizable_Plotting_Function, echo=FALSE}
##
plot.gtrends.silent <- function(x, ...) {
  df <- x$interest_over_time
  df$date <- as.Date(df$date)
  df$hits <- if(typeof(df$hits) == 'character'){
    as.numeric(gsub('<','',df$hits))
    } else {
    df$hits
    }
 
  df$legend <- paste(df$keyword, " (", df$geo, ")", sep = "")
 
  p <- ggplot(df, aes_string(x = "date", y = "hits", color = "legend")) +
    geom_line() +
    xlab("Date") +
    ylab("Search Hits") +
    ggtitle("Interest over time") +
    theme_bw() +
    theme(legend.title = element_blank())
 
  invisible(p)
}
```

Methodology per Dave Tang, [Visualising Google Trends Results with R](https://davetang.org/muse/2018/12/31/visualising-google-trends-results-with-r/)

```{r single_hit_yaxis_interest_plot, echo=FALSE}

fire_and_climate_change_search_terms <- c(
  "fire",
  "climate change",
  "insurance"
)

CA_all_fire_trends <- gtrends(keyword = fire_and_climate_change_search_terms,
        geo = "CA") #time = "today+5-y" Last five years (default)

CA_fire_plot <- plot.gtrends.silent(CA_all_fire_trends)
 
CA_fire_plot +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  theme(legend.position = "bottom")
```

Methodology per Surbhi Tyagi, [Getting Google Trends Data Using gtrendsR](http://rstudio-pubs-static.s3.amazonaws.com/520533_f25fa191afa7476a995322a4393710ce.html)

```{r relative_yaxis_hit_plot, echo=FALSE}

indiv_fire_trend <- gtrends(keyword = "fire",
        geo = "CA") #time = "today+5-y" Last five years (default)

indiv_cc_trend <- gtrends(keyword = "climate change",
        geo = "CA") #time = "today+5-y" Last five years (default)

date <- data.frame(indiv_fire_trend[["interest_over_time"]][["date"]])
fire_hits <- data.frame(indiv_fire_trend[["interest_over_time"]][["hits"]])
cc_hits <- data.frame(indiv_cc_trend[["interest_over_time"]][["hits"]])

temp_fire_df <- cbind(date, fire_hits)
temp_cc_df <- cbind(date, cc_hits)

names(temp_fire_df)[1] <- "date"
names(temp_fire_df)[2] <- "fire_hits"

names(temp_cc_df)[1] <- "date"
names(temp_cc_df)[2] <- "cc_hits"

temp_joined <- inner_join(temp_fire_df, temp_cc_df, by = NULL) #def. by="date"
view(temp_joined)

graph_df <- temp_joined %>% 
  pivot_longer(!date, names_to = "type", values_to = "hits")

ggplot(data = graph_df) + 
  aes(x = date, y = hits, color = type) + 
  geom_line() + 
  scale_colour_viridis_d(option  = "viridis") + 
  labs(title = "Interest by Date", 
       x = "Date", 
       y = "Relative Interest") + 
  #scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  theme_minimal() + 
  theme(legend.position = 'bottom')
```

---

# Research Question 2

**Does interest in climate change increase regionally after the occurrence of a **
**natural disaster?**

By virtue of research question 1, it is reasonable to assume a relationship 
between natural disasters - physical events that shape critical aspects of the
human experience - and Google search trends for these events. When FEMA declares 
an emergency, for a hurricane, wildfire, flood, tornado or other event, people 
who reside in the affected region want to know what is going on. This includes 
searching the internet for more information about the nature of the emergency, 
advice regarding how to protect their families and property, what to expect, and 
how to prepare. The Google trends hits value associated with the disastr's name 
thus fluctuates dramatically in the days leading up to, during, and after a 
natural disaster impacts a community.

The next logical question to pose - at least from the perspective of this 
analysis - is whether or not humans have a tendency to search for topics like 
climate change after the occurrence of a local natural disaster. One of the most
noticeable and consequential results of increased global warming is that extreme
weather events, like storms, fires and floods, become more severe and can occur 
more frequently. This is often used as a rallying-cry by climate activists. 
Without timely action to minimize greenhouse gas emissions and protect our 
natural resources, there will likely be irreversible damage to the planet. 

## Florida Hurricanes since `r date_5y`

A list of all FEMA declarations for hurricanes in Florida occurring in the last 
5 years.

```{r FL_Hurricanes-events, echo=FALSE}
FL_Hurricanes <- FEMA_Declarations %>% 
  filter(incident_type == "Hurricane", incident_begin_date >= date_5y, state == "FL") %>% 
  select(declaration_title, declaration_day) %>% 
  distinct(declaration_title, .keep_all = TRUE)

FL_Hurricanes
```

```{r FL_Hurricanes-query}

# set parameters for FL_Hurricanes query below
num_days <- 14
State_Disaster <- FL_Hurricanes
location <- "US-FL"
search_term <- "climate change"

# vectorize events & dates
events <- as.vector(State_Disaster$declaration_title)
dates <- as.vector(State_Disaster$declaration_day)

# create vector of length length(events)*num_days with event names for cbind()
EventsVec <- vector()
for (j in 1:length(events)) {
  event <- events[j]
  for(k in 1:(num_days+1)) {
    EventsVec <- append(EventsVec, event)
  }
}


# initialize empty data frame
finaldata <- as.data.frame(NULL)

# for loop iterates over all events in events vector
for (i in 1:length(events)) {
  
  # create comb_time time boundaries for gtrends query
  start_date <- ymd(dates[i])
  end_date <- ymd(dates[i]) + num_days
  comb_time <- paste(start_date, end_date)

  # execute gtrends query for keyword = list_of_weather_events[i]
  #
  event_trend <- gtrends(keyword = search_term, #events[i]
                     geo = location, time = comb_time)
  
  # weather_event_gtrends[1] object tracks interest_over_time
  all_time_interest <- as.data.frame(event_trend[1])
  finaldata <- rbind(finaldata, all_time_interest)

  print(i)
  Sys.sleep(5)
}

# column bind df with vector of natural disaster names (EventsVec)
finaldata <- cbind(event_name=EventsVec, finaldata)

# set finaldata pointer to FLHurr_Hits (specific variable)
FLHurr_Hits <- finaldata
# view data frame
FLHurr_Hits
```

```{r FL_Hurricanes-hits-df}

# set df as FLHurr_Hits for subsequent df transformation
df <- FLHurr_Hits
num_days <- 14

x <- df %>%
  mutate(event_date = gsub( " .*$", "", interest_over_time.time ),
         hits_date = gsub( " .*$", "", interest_over_time.date )
         )
  
x <- x %>% 
  mutate(event_date = as.Date(event_date),
         hits_date = as.Date(hits_date),
         keyword = interest_over_time.keyword,
         hits = interest_over_time.hits) %>% 
  select(event_name, event_date, keyword, hits_date, hits) 

relative_date_mod <- x %>% 
  mutate(rel_date = hits_date - event_date)

# ggplot ready df (pivot_longer effectively done)
FLHurr_graph_df <- relative_date_mod
```

# method to get 1 row for each event
hits_df <- relative_date_mod %>% 
  pivot_wider(names_from = rel_date, values_from = hits)

coalesce_by_column <- function(df) {
  return(coalesce(df[1:(num_days)]))
}

hits_df %>%
  group_by(event_name) %>% #event_name
  summarise_all(coalesce_by_column)

FLHurr_output_df <- hits_df
FLHurr_output_df

`coalesce_by_column` function [description and source](https://stackoverflow.com/questions/45515218/combine-rows-in-data-frame-containing-na-to-make-complete-row)

```{r FLHurr_CCInterest-plot}
ggplot(data = FLHurr_graph_df, aes(x = rel_date, y = hits, color = event_name)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "FL Hurricanes Affects on Climate Change Sentiment",
       subtitle = "Google Trends after FEMA Declaration",
       x = "Number of Days after FEMA Declaration",
       y = "Google Trends Hits Volume",
       color = "Natural Disaster") +
  theme(legend.position = "right")
```

```{r}
#State_Disaster <- FL_Hurricanes
# add a grouping variable (or many!)
 FLHurr_slopes <- FLHurr_graph_df %>%
  #mutate(group5 = rep(1:10, each = 5)) %>%
  group_by(event_name) %>%
  mutate(
    slope = round(lm(hits ~ rel_date)$coefficients[2], 2),
    significance = summary(lm(hits ~ rel_date))$coefficients[2, 4],
    x = mean(rel_date),   # x coordinate for slope label
    y = mean(hits)     # y coordinate for slope label
  )
FLHurr_slopes %>% 
  select(event_name, event_date, slope, significance) %>% 
  distinct()  %>%
  filter(significance < .2)
```


```{r FLHurr_CCMono-plot}
ggplot(data = FLHurr_graph_df, aes(x = rel_date, y = hits)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "FL Hurricanes Affects on Climate Change Sentiment",
       subtitle = "Google Trends after FEMA Declaration",
       y = "Google Trends Hits Volume") +
  scale_x_continuous("Number of Days after FEMA Declaration", labels = as.character(xticks), breaks = xticks) + 
  theme_bw()
```

## California Wildfires since `r date_5y`

A list of all FEMA declarations for wildfires in California occurring in the 
last 5 years.

```{r CA_Fires-events, echo=FALSE}
CA_Fires <- FEMA_Declarations %>% 
  filter(incident_type == "Fire", incident_begin_date >= date_5y, state == "CA") %>% 
  select(declaration_title, declaration_day) %>% 
  distinct(declaration_title, .keep_all = TRUE)
  
CA_Fires
```

```{r CA_Fires-query}

# set parameters for CA_Fires query below
num_days <- 14
State_Disaster <- CA_Fires
location <- "US-CA"
search_term <- "climate change"

# vectorize events & dates
events <- as.vector(State_Disaster$declaration_title)
dates <- as.vector(State_Disaster$declaration_day)

# create vector of length length(events)*num_days with event names for cbind()
EventsVec <- vector()
for (j in 1:length(events)) {
  event <- events[j]
  for(k in 1:(num_days+1)) {
    EventsVec <- append(EventsVec, event)
  }
}

# initialize empty data frame
finaldata <- as.data.frame(NULL)
# for loop iterates over all events in events vector
for (i in 1:length(events)) {
  
  # create comb_time time boundaries for gtrends query
  start_date <- ymd(dates[i])
  end_date <- ymd(dates[i]) + num_days
  comb_time <- paste(start_date, end_date)

  # execute gtrends query for keyword = list_of_weather_events[i]
  event_trend <- gtrends(keyword = search_term, #events[i]
                     geo = location, time = comb_time)
  
  # weather_event_gtrends[1] object tracks interest_over_time
  all_time_interest <- as.data.frame(event_trend[1])
  finaldata <- rbind(finaldata, all_time_interest)
  print(i)
  Sys.sleep(5)
}

# column bind df with vector of natural disaster names (EventsVec)
finaldata <- cbind(event_name=EventsVec, finaldata)

# set finaldata pointer to CAFire_Hits (specific variable)
CAFire_Hits <- finaldata
# view data frame
CAFire_Hits
```

```{r CA_Fires-hits-df}

# set df as CAFire_Hits for subsequent df transformation
df <- CAFire_Hits

x <- df %>%
  mutate(event_date = gsub( " .*$", "", interest_over_time.time ),
         hits_date = gsub( " .*$", "", interest_over_time.date )
         )
  
x <- x %>% 
  mutate(event_date = as.Date(event_date),
         hits_date = as.Date(hits_date),
         keyword = interest_over_time.keyword,
         hits = interest_over_time.hits) %>% 
  select(event_name, event_date, keyword, hits_date, hits) 

relative_date_mod <- x %>% 
  mutate(rel_date = hits_date - event_date)

# ggplot ready df (pivot_longer effectively done)
CAFire_graph_df <- relative_date_mod
```


##TAKEOUT
hits_df <- relative_date_mod %>% 
  pivot_wider(names_from = rel_date, values_from = hits)

coalesce_by_column <- function(df) {
  return(coalesce(df[1 : num_days]))
}

hits_df %>%
  group_by(keyword) %>%
  summarise_all(coalesce_by_column)
  
CAFire_output_df <- hits_df

```{r CAFire_CCInterest-plot}
ggplot(data = CAFire_graph_df, aes(x = rel_date, y = hits, color = event_name)) +
  geom_point(show.legend = FALSE) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "CA Fires Affects on Climate Change Sentiment",
       subtitle = "Google Trends after FEMA Declaration",
       x = "Number of Days after FEMA Declaration",
       y = "Google Trends Hits Volume") +
  theme(legend.position = "none")
```

```{r CAFire_CCMono-plot}
xticks <- 0:num_days
ggplot(data = CAFire_graph_df, aes(x = rel_date, y = hits)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "CA Fires Affects on Climate Change Sentiment",
       subtitle = "Google Trends after FEMA Declaration",
       y = "Google Trends Hits Volume") +
  scale_x_continuous("Number of Days after FEMA Declaration", labels = as.character(xticks), breaks = xticks) + 
  theme_bw()
```

```{r single_hit_yaxis_interest_plot, echo=FALSE}

fire_and_climate_change_search_terms <- c(
  "fire",
  "climate change",
  "insurance"
)

CA_all_fire_trends <- gtrends(keyword = fire_and_climate_change_search_terms,
        geo = "CA") #time = "today+5-y" Last five years (default)

CA_fire_plot <- plot.gtrends.silent(CA_all_fire_trends)
 
CA_fire_plot +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  theme(legend.position = "bottom")
```

Methodology per Surbhi Tyagi, [Getting Google Trends Data Using gtrendsR](http://rstudio-pubs-static.s3.amazonaws.com/520533_f25fa191afa7476a995322a4393710ce.html)

```{r relative_yaxis_hit_plot, echo=FALSE}

indiv_fire_trend <- gtrends(keyword = "fire",
        geo = "CA") #time = "today+5-y" Last five years (default)

indiv_cc_trend <- gtrends(keyword = "climate change",
        geo = "CA") #time = "today+5-y" Last five years (default)

date <- data.frame(indiv_fire_trend[["interest_over_time"]][["date"]])
fire_hits <- data.frame(indiv_fire_trend[["interest_over_time"]][["hits"]])
cc_hits <- data.frame(indiv_cc_trend[["interest_over_time"]][["hits"]])

temp_fire_df <- cbind(date, fire_hits)
temp_cc_df <- cbind(date, cc_hits)

names(temp_fire_df)[1] <- "date"
names(temp_fire_df)[2] <- "fire_hits"

names(temp_cc_df)[1] <- "date"
names(temp_cc_df)[2] <- "cc_hits"

temp_joined <- inner_join(temp_fire_df, temp_cc_df, by = NULL) #def. by="date"
view(temp_joined)

graph_df <- temp_joined %>% 
  pivot_longer(!date, names_to = "type", values_to = "hits")

ggplot(data = graph_df) + 
  aes(x = date, y = hits, color = type) + 
  geom_line() + 
  scale_colour_viridis_d(option  = "viridis") + 
  labs(title = "Interest by Date", 
       x = "Date", 
       y = "Relative Interest") + 
  #scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  theme_minimal() + 
  theme(legend.position = 'bottom')
```

---

# Research Question 3

**Do people modify their Google search behavior and become more **
**environmentally conscious after a natural disaster?**


---

# Bibliography

## Existing Scholarship on Climate Change Sentiment Analysis

1. CHOI, H. and VARIAN, H. (2012), Predicting the Present with Google Trends. 
Economic Record, 88: 2-9. https://doi.org/10.1111/j.1475-4932.2012.00809.x

2. "What do Google searches tell us about our climate change fears?" 
NewsRx Health & Science, 3 Aug. 2014, p. 177. Gale Academic OneFile, https://link.gale.com/apps/doc/A378494502/AONE?u=duke_perkins&sid=AONE&xid=7d6f5991. Accessed 16 Nov. 2020.

3. 

## Technical Sources for Programming & Visualizations

1. https://cran.r-project.org/web/packages/gtrendsR/gtrendsR.pdf

2. https://davetang.org/muse/2018/12/31/visualising-google-trends-results-with-r/

3. http://rstudio-pubs-static.s3.amazonaws.com/520533_f25fa191afa7476a995322a4393710ce.html

4. 
